Progress with the NOSAMS

### 2021-05-02 doc/SNEAP2021/SNEAP2021_HGIS_presentation/SNEAP2021_HGIS_presentation.Rmd
Another round of edits.

.pull-right[
![Salazar, 2013](images/salazar_hgis_target_flow.png)
.right[Salazar, 2013]
]

@@

.pull-left[
* Introduce gas to back of cathode
* Rest is gas handling
<br>
<br>

```{r hgis-arm, echo=FALSE, fig.align='center', out.width="80%"}
knitr::include_graphics("images/hgis_arm_source.jpeg")
```
]

@@

.pull-right[
![HGIS Target](images/target_cutaway.png)
NEC hybrid target cutaway
]

@@

### Results
* ~10μA <sup>12</sup>C<sup>-</sup>
* <1% precision (SE of repeat measurements)
* 1% blank
* 1-3% total efficiency
]

.pull-right[
```{r cur-flow, fig.height=6, dev='svg'}
read.csv(here("data/14Nov2017LPtest.csv"), skip = 4) %>%
  mutate(flow = flowcalc(supply/10, 2E-5, 1.49E-5, 1.58)) %>% 
  ggplot(aes(flow, current)) +
  geom_smooth(se = FALSE, span = 0.7, color = dark_blue) + 
  geom_point() +
  labs(title = "Current vs gas flow",
          x = bquote('CO'[2]~'Flow (μl/min)'), 
          y = "12C current (μA)") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff")
```
]

@@

Increase flow by
* increasing pCO<sub>2</sub>
* reducing volume

Gas mixtures - partial pressure
* adding another gas increases pressure and total flow
* reduces pCO<sub>2</sub>
* CO<sub>2</sub> remains constant? No- viscosity
]

@@

.pull-right.center[
<br><br><br>
<img src="images/dual_bellows.png" width="80%" />
]

@@

* VG Prism dual bellows
* LabVIEW interface
* 2 x bellows
* Cryotraps
* Manifold capable

![IRMS inlet diagram](images/HGIS_dual_bellows_drawing.svg)
]

@@

P and V are held constant, so only mixture controls pCO<sub>2</sub>

Use helium to displace CO<sub>2</sub> from vials into split.

Capillary to source controls gas flow

Open inlets maintain constant atmospheric pressure, producing constant<sup>1</sup> flow to source

![open split diagram](images/HGIS_open_split_drawing.svg)

.footnote[ [1] flow varies by composition due to differing viscosity]
]

.pull-right.center[
<br><br><br>
<img src="images/gas_control.jpg" width="75%" />
]

@@

* Evacuate vial
* Add acid to carbonate
* Fills vial with CO<sub>2</sub> at 1 atm

@@

* Pre-sputter target
* Displace sample 
* Measurement
]

.pull-right[
.center[![gilson autosampler](images/gilson.png)]
]

@@

* No current suppresion at high CO<sub>2</sub>
* Current stable until 40 min
* Corresponds to 2mL CO<sub>2</sub>

Balance flow rates to maximize stable currents and minimize sample size
]

.pull-right[
```{r cur-time-os, fig.height=8, dev='svg'}
data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
  filter(Pos == 5) %>% 
  mutate(time = cum_acqtime/60,
         co2flow = concCO2(time, r = 244) * 30) #30ul/min

@@

# Currents and flows

.pull-left[
* Note similarity to first tests
* Flat top means stable currents/ratios for long enough to measure
* Sample size can be reduced to 2mL
* Further reduction with increased flow to source
]

.pull-right[
```{r cur-flow-os, fig.height=6, dev='svg'}
ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE, color = medium_blue) +
  geom_point() +
  labs(title = "Vial dillution current",
       subtitle = "250μL/min displacement, 30μl/min delivery",
       x = bquote('CO'[2]~'Flow (μl/min)'), 
       y = "12C current (μA)") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff")
```
]

@@

0.1μA blank current with 10μA total means 1% blank

@@

.pull-left[
If standards, blanks, and unknowns have same current:

linear fit used to correct

@@

#### Mean Fm of blanks:
`r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`

@@

* Currents > 2μA
* Note current dependence

@@

Source differences
* one target in source at a time - lower blank
* gas supplied via conical joint in side of target

Inlet differences
* MICADAS uses closed inlet
* CO<sub>2</sub> from carbonates transferred onto resin trap in stream of helium
* Sample introduced by adding helium and adjusting flow using variable volume
]

.pull-right[![](images/MICADAS_HGIS_interface.jpeg)]

@@

NOSAMS has developed a successful method for analysing carbonates with an open-split interface delivering gas to a NEC MCSNICS gas-accepting Cesium sputter ion source.

Current goals:
* finish characterization of the method
* Comparison to graphite with test samples
* Improve accuracy
* Better understand method blank
* Reduce sample size

>>>>

.pull-right[
![Salazar, 2013](images/salazar_hgis_target_flow.png)
.right[Gas density in HGIS Cathode

(Salazar, 2013)]
]

@@

* Pneumatic arm brings gas to source
* Introduce gas to back of cathode
* The rest is gas handling
<br>
<br>

.center[<img src="images/hgis_arm_source.jpeg" width="80%" />]
.right[MCSNICS source with gas arm]
]

@@

.pull-right[
![HGIS Target](images/target_cutaway.png)
.right[NEC hybrid target cutaway]
]

@@

### Results
* ~8 μA <sup>12</sup>C<sup>-</sup>
* <1% precision (SE of repeat measurements)
* 1% blank
* 1-3% total efficiency
]

.pull-right[
```{r cur-flow, fig.height=6, dev='svg'}
read.csv(here("data/14Nov2017LPtest.csv"), skip = 4) %>%
  mutate(flow = flowcalc(supply/10, 2E-5, 1.49E-5, 1.58)) %>% 
  ggplot(aes(flow, current)) +
  geom_smooth(se = FALSE, span = 0.7, color = dark_blue) + 
  geom_point() +
  labs(title = "Ion current and gas flow",
          x = bquote('CO'[2]~'Flow (μl/min)'), 
          y = "12C current (μA)") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff")
```
]

@@

Increase CO<sub>2</sub> flow by increasing pCO<sub>2</sub>:
* Add CO<sub>2</sub> 
* Reduce volume

Gas mixtures - helium dilution:
* Increase pressure and total flow
* Reduce pCO<sub>2</sub>
* CO<sub>2</sub> flow remains constant? No - viscosity
]

@@

.pull-right.center[
<br><br><br>
<img src="images/dual_bellows.png" width="80%" />

.right[Dual-bellows closed inlet]
]

@@

* VG Prism inlet
* Bellows for volume control
* Cryotraps to move and isolate CO<sub>2</sub>
* LabVIEW control
* Manifold capable

![IRMS inlet diagram](images/HGIS_dual_bellows_drawing.svg)
.right[Closed inlet schematic]
]

@@

* Pressure is held constant at 1 atm
* Gas mixture controls pCO<sub>2</sub>
* Helium displaces CO<sub>2</sub> from vials into split
* Capillary to source controls gas flow

Open inlets maintain constant atmospheric pressure, producing constant<sup>1</sup> flow to source.

![open split diagram](images/HGIS_open_split_drawing.svg)
]

.pull-right.center[
<br><br>
<img src="images/gas_control.jpg" width="75%" />
.right[Open inlet setup]
]

.footnote[[1] flow varies by composition due to differing viscosity]

@@

* Weigh carbonate into vial
* Evacuate vial
* Add acid to carbonate
* Reaction fills vial with CO<sub>2</sub> at 1 atm

@@

* Pre-sputter target
* Displace sample from vial to split with He 
* Measure ratios for 15-30 min.
]

.pull-right[
.center[![gilson autosampler](images/gilson.png)]
.right[Automated gas handling system]
]

@@

* No current suppression at high pCO<sub>2</sub>
* Current stable until 40 min
* 40 min ≈ 2 mL CO<sub>2</sub>
]

.pull-right[
<br>
<br>
```{r cur-time-os, fig.height=7, dev='svg'}
data <- get_hgis_data(here("data/USAMS040121R.txt"), as.Date("2021-04-02")) %>% 
  filter(Pos == 5) %>% 
  mutate(time = cum_acqtime/60,
         co2flow = concCO2(time, r = 244) * 30) #30ul/min

@@

* Note similarity to first tests
* Flat top means stable currents/ratios for long enough to measure
* Sample size can be reduced to 2 mL
* Further reduction with increased flow to source

### Balance flow rates to maximize stable currents and minimize sample size
]

.pull-right[
<br>
<br>
<br>
```{r cur-flow-os, fig.height=6, dev='svg'}
ggplot(data, aes(co2flow, he12C)) +
  geom_smooth(span = .3, se=FALSE, color = medium_blue) +
  geom_point() +
  labs(title = "Vial dillution current",
       subtitle = "250μL/min displacement, 30μl/min delivery",
       x = bquote('CO'[2]~'Flow (μl/min)'), 
       y = "12C current (μA)") +
  theme_classic() +
  theme_xaringan(title_font_size = 20,
                 text_font_size = 16,
                 background_color = "#ffffff")
```
]

@@

0.1 μA blank current with 10 μA total = 1% blank

@@

.pull-left[
If **standards**, **blanks**, and **unknowns** have the **same** blank and current:

### Linear fit used to correct:

@@

### Mean Fm of blanks: `r carb_data %>% filter(fm_consensus < 0.1 & Cur > 2) %>% pull(mean) %>% mean() %>% format(digits = 3)`

@@

* Currents > 2 μA
* Note current dependence

@@

#### Source differences
* one target in source at a time - lower blank
* gas supplied via conical joint in side of target

#### Inlet differences
* MICADAS uses closed inlet
* CO<sub>2</sub> from carbonates transferred onto resin trap in stream of helium
* Sample introduced by adding helium and adjusting flow using variable volume
]

.pull-right[
![](images/MICADAS_HGIS_interface.jpeg)
.right[MICADAS Gas Interface]
]

@@

NOSAMS has developed a successful method for analyzing radiocarbon in carbonates with an open-split interface delivering CO<sub>2</sub> gas to a NEC MCSNICS gas-accepting cesium sputter ion source.

Current goals:
* Finish characterization of the method
* Compare to graphite with test samples
* Improve accuracy
* Better understand method blank
* Reduce sample size
* Automation of data acquisition and reduction
